On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-10-01 paper.tex
Moved everything

@@

@@

In the sequel we will also need a {\it relative Steinberg symbol}, i.\,e. a map
\begin{equation*} \{-, -\}'_\alpha \colon R^* \times (1 + I)^* \to \St(\Phi, R, I) \end{equation*}
such that $ \ext{I} \{u, v\}'_\alpha = \{u, v\}_\alpha$. It is not hard to verify that the image under $\mu_I$ of the element
\begin{equation*} f(\alpha, t, u)= y_\alpha(t(u-1)) \cdot y_{-\alpha}(-t^{-1}(u^{-1}-1))^{x_\alpha(-t)} \cdot y_{\alpha}(t(u-1))^{w_\alpha(-t)}, \end{equation*}
equals $w_\alpha(tu) \cdot w_\alpha(-t)$. This allows us to set $\{t, u \}'_\alpha := f(\alpha, t, u) f(\alpha, 1, u)^{-1}.$
Now, since $\mu_I\{ t, u \}'_\alpha = w_\alpha(tu) w_\alpha(-t) w_\alpha(1) w_\alpha(-u) = \{t, u\}_\alpha,$ 
 we have defined the desired symbol.
\begin{rem}
 Our relative Steinberg symbols differs from the one defined in~\cite{Ste73} in that
  it takes values in $\St(\Phi, R, I)$ rather than in $\St(\Phi, R)$. %Elaborate on this...
\end{rem}

\subsection{Another presentation of linear Steinberg groups} \label{sec:another-presentation}
%TODO: Formulate why do we write what we write
Throughout this subsection $R$ denotes a commutative ring.

For $u \in R^n$ we denote by $D(u)$ the subset of $R^n$ consisting of all $v$ having at least two zero entries and satisfying $u^tv = 0$.
Recall from 3.2 of~\cite{Ka77} that for every $u, v, w \in R[t]^n$ such that $u^t \cdot v = 0$ there
 is a decomposition of $(w^t u) \cdot v$ into a sum of elements of $D(u)$, called {\it canonical decomposition}:

\begin{equation} \label{eq:canonical} (w^tu) \cdot v=\sum_{i<j}u_{ij} c_{ij}(v, w),\text{ where $u_{ij}=e_iu_j-e_ju_i,$ $c_{ij}(v, w)=v_iw_j-v_jw_i.$}\end{equation}
Let $u, v \in R^n$ be such that $u^t v = 0$ and assume, moreover, that either $u$ or $v$ has at least one zero entry.
Recall from 3.8, 3.10 of~\cite{Ka77} that under these assumptions one can define certain element
 $x(u, v)$ of $\St(n, R)$ such that $\phi(x(u, v)) = t(u, v) = 1 + uv^t$.
The following properties of $x(u, v)$ are standard (cf.~\cite[Lemma~1.1]{Tu83}).
\begin{lemma} \label{lem:xsmall-properties}
\begin{lemlist}
 \item \label{item:xsmall-scalar} If $v$ or $w$ has at least two zero entries, then 
 \begin{equation}\nonumber x(v, wa) = x(va, w).\end{equation} %TODO: What is ''a``?
 \item \label{item:xsmall-additivity} If $w_1$ and $w_2$ have at least one common zero entry then
 \begin{equation}\nonumber x(v, w_1)x(v, w_2) = x(v, w_1+w_2).\end{equation}
 \item \label{item:xsmall-commute} If $v$, $v'$ are simultaneously orthogonal to $w$ and $w'$ and the elements $w, w'$
  both have at least two zero entries then $x(v, w)$ and $x(v', w')$ commute.
\end{lemlist}
\end{lemma}

Now let $I$ be a splitting ideal of $R$.
Let $u \in E(n, R)e_1$, $v \in I^n$ be vectors such that $u^tv = 0$.
Recall from \S~4 of~\cite{LS17} that one can define the following elements of $\St(n, R, I) \leq \St(n, R)$:
\begin{equation} \label{eq:sigma-definition}
 F(u, v) = \prod\limits_{i=1}^r x(u,  v_i),\ \
 S(v, u) = \prod\limits_{i=1}^r x( v_i, u).
\end{equation}
Here $\{v_r\}$, $r=1,\ldots,N$ is any collection of elements of $I^n \cap D(u)$ such that $\sum_{r=1}^N v_r = v$.
Since $u$ is unimodular, such a collection always exists by~\eqref{eq:canonical}, moreover $F(u, v)$ and $S(u, v)$ do not depend on its choice. %TODO: Give reference

It is possible to give a presentation of the relative Steinberg group $\St(n, R, I)$ formulated in terms of elements $F(u, v)$, $S(v, u)$.
\begin{prop}[\text{\cite[Proposition 3.10]{LS17}}] \label{prop:rel-presentation}
 For any $n\geq 4$ the group $\St(n,\,R,\,I)$ can be presented by means of generators $F(u,\,v)$, $S(v,\,u)$ (where $u\in E(n,\,R)e_1,$ and $v\in I^n$ are such that $u^tv=0$) subject to the following relations:
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)^{-1}=F(t(u,\,v)u',\,t(v,\,u)^{-1} v'), \label{conj3} \\
&F(ge_1,\,g^*e_2a)=S(ge_1a,\,g^*e_2),\ \text{for all}\ a\in I,\, g \in E(n, R).
\end{align}
\end{prop}
In the above statement $g^*$ denotes the contragradient matrix, i.\,e. $g^* = {g^{-1}}^t$ and $e_i$ is the standard basis vector of $R^n$.
\begin{rem}
Using essentially the same reasoning as in~\cite[Proposition~8]{S15} it is not hard to show that the above presentation still holds for $\St(\Phi, R, I)$ without the assumption that $I$ is splitting.
\end{rem}

@@

Indeed, define $\sigma_1$ on $N_{0}=\St(n, A[t], tA[t])$ by the following identities: 
\begin{equation}
 \sigma_1(F(u, v)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot t u, \delta_1 \cdot t^{-1} v_i),\ \
 \sigma_1(S(v, u)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot v_i, \delta_1 \cdot u).
\end{equation}

>>>>

For $g\in \GL(n, R)$ we denote by $g^*$ the contragradient matrix, i.\,e. $g^* = {g^{-1}}^t$.
We denote by $e_i$ the standard basis vector of $R^n$. For a pair of vectors $u, v \in R^n$ such that $u^tv = 0$ we denote by
$t(u, v)$ the rank $1$ matrix $e + uv^t$. Notice that $t(u, v)$ lies in $\E(n, R)$ provided $u$ or $v$ is unimodular. %TODO: Reference

%TODO: Introduce notation for GL(n, R); E(n, R)

@@

\subsection{Another presentation of linear Steinberg groups} \label{sec:another-presentation}
Throughout this subsection $R$ denotes a commutative ring.

For $u \in R^n$ we denote by $D(u)$ the subset of $R^n$ consisting of all vectors $v$ which are orthogonal to $u$ (i.\,e. $u^tv = 0$) and have at least two zero entries.
Recall from 3.2 of~\cite{Ka77} that for every $u, v, w \in R^n$ such that $u^t v = 0$ there
 is a decomposition of $(w^t u) \cdot v$ into a sum of elements of $D(u)$, called {\it canonical decomposition}:
\begin{equation} \label{eq:canonical} (w^tu) \cdot v=\sum_{i<j}u_{ij} c_{ij}(v, w),\end{equation}
where $u_{ij} =e_iu_j-e_ju_i \in D(u)$ and $c_{ij}(v, w) =v_iw_j-v_jw_i \in R$.
Let $u, v \in R^n$ be such that $u^t v = 0$ and assume, moreover, that either $u$ or $v$ has at least one zero entry.
Recall from 3.8, 3.10 of~\cite{Ka77} that under these assumptions one can define certain element
 $x(u, v)$ of $\St(n, R)$ such that $\phi(x(u, v)) = t(u, v) = 1 + uv^t$.
Let us recall the standard properties of $x(u, v)$ (cf.~\cite[Lemma~1.1]{Tu83}).
\begin{itemize}
 \item If $v$ or $w$ has at least two zero entries, then
  \begin{equation} \label{item:xsmall-scalar} x(v, wa) = x(va, w), \text{ for $a\in R$} \end{equation}
 \item If $w_1$ and $w_2$ have at least one common zero entry then
 \begin{equation} \label{item:xsmall-additivity} x(v, w_1) \cdot x(v, w_2) = x(v, w_1+w_2).\end{equation}
 \item If $v$, $v'$ are simultaneously orthogonal to $w$ and $w'$ and the elements $w, w'$ both have at least two zero entries then 
  \begin{equation} \label{item:xsmall-commute} [x(v, w),\ x(v', w')] = 1 \end{equation}
\end{itemize}
\begin{comment}
Now let $I$ be a splitting ideal of $R$.
Let $u \in E(n, R)e_1$, $v \in I^n$ be vectors such that $u^tv = 0$.
Recall from \S~4 of~\cite{LS17} that one can define the following elements of $\St(n, R, I) \leq \St(n, R)$:
\begin{equation} \label{eq:sigma-definition}
 F(u, v) = \prod\limits_{i=1}^r x(u,  v_i),\ \
 S(v, u) = \prod\limits_{i=1}^r x( v_i, u).
\end{equation}
Here $\{v_r\}$, $r=1,\ldots,N$ is any collection of elements of $I^n \cap D(u)$ such that $\sum_{r=1}^N v_r = v$.
Since $u$ is unimodular, such a collection always exists by~\eqref{eq:canonical}, moreover $F(u, v)$ and $S(u, v)$ do not depend on its choice. %TODO: Give reference
\end{comment}

@@

In the sequel we will also need a {\it relative Steinberg symbol}, i.\,e. a map
\begin{equation*} \{-, -\}'_\alpha \colon R^* \times (1 + I)^* \to \St(\Phi, R, I) \end{equation*}
such that $ \ext{I} \{u, v\}'_\alpha = \{u, v\}_\alpha$. It is not hard to verify that the image under $\mu_I$ of the element
\begin{equation*} f(\alpha, t, u)= y_\alpha(t(u-1)) \cdot y_{-\alpha}(-t^{-1}(u^{-1}-1))^{x_\alpha(-t)} \cdot y_{\alpha}(t(u-1))^{w_\alpha(-t)}, \end{equation*}
equals $w_\alpha(tu) \cdot w_\alpha(-t)$. This allows us to set $\{t, u \}'_\alpha := f(\alpha, t, u) f(\alpha, 1, u)^{-1}.$
Now, since $\mu_I\{ t, u \}'_\alpha = w_\alpha(tu) w_\alpha(-t) w_\alpha(1) w_\alpha(-u) = \{t, u\}_\alpha,$ 
 we have defined the desired symbol.
\begin{rem}
 Our relative Steinberg symbols differs from the one defined in~\cite{Ste73} in that
  it takes values in $\St(\Phi, R, I)$ rather than $\St(\Phi, R)$. %Elaborate on this...
\end{rem}

In the sequel we will need an explicit presentation of the relative linear Steinberg group by means of generators and relations.
The presentation we formulate below should be thought of as an ''improved`` version of the presentation of the relative linear Steinberg group
 given by Tulenbaev in~\cite[Definition~1.5]{Tu83} (cf. also~\cite[Proposition~3.2]{LS17}).
We refer the reader to~\cite[Section~3]{LS17} for the discussion why this presentation is better than Tulenbaev's one
 (the short answer is: it remains useful in the case $\Phi = \rA_3$).

\begin{prop}[\text{\cite[Proposition 3.10]{LS17}}] \label{prop:rel-presentation}
 Assume that $I$ is a splitting ideal of a commutative ring $R$. 
 Then for any $\ell\geq 3$ the group $\St(\rA_\ell,\,R,\,I)$ can be presented by means of two families of generators $F(u,\,v)$, $S(v,\,u)$
  (where $u\in \E(n,\,R)e_1,$ and $v\in I^n$ are such that $u^tv=0$) subject to the following relations:
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)^{-1}=F(t(u,\,v)u',\,t(v,\,u)^{-1} v'), \label{conj3} \\
&F(ge_1,\,g^*e_2a)=S(ge_1a,\,g^*e_2),\ \text{for all}\ a\in I,\, g \in E(n, R).
\end{align}
\end{prop}
\begin{rem}
Although we will not need this, using essentially the same reasoning as in~\cite[Proposition~8]{S15} it is not hard to show that 
$\St(\rA_\ell, R, I)$ still admits the above presentation without the assumption that $I$ is splitting.
\end{rem}

@@

%TODO: Pick a decomposition  $v = \sum v_i$
Indeed, for $u \in E(n, A[t])$ and $v \in tA[t]$ define $\sigma_1$ on $N_{0}=\St(n, A[t], tA[t])$ by the following identities: 
\begin{equation}
 \sigma_1(F(u, v)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot t u, \delta_1 \cdot t^{-1} v_i),\ \
 \sigma_1(S(v, u)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot v_i, \delta_1 \cdot u).
\end{equation}
